---
title: "TVE 16S Diversity Analyses"
author: "Nicola Kriefall, updated by Hannah Aichelman"
date: "2/24/2022"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
---

```{r setup, include=FALSE}
#work computer 
setwd("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/")
```

# Setup

## Packages

```{r packages}
library(ggplot2)
library(cowplot)
library(phyloseq)
library(car)
library(ggpubr)
library(vegan)
library(dada2)
library(tidyverse)
```

## Read in data

```{r read in data}
samdf <- read.csv("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/samdf_noL3.csv",header=TRUE) %>%
  select(-X)
row.names(samdf) <- samdf$frag
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/taxa2.Rdata")
```

## Phyloseq objects

```{r making phyloseq objects, eval=FALSE}
#rarefied version
ps.rare = readRDS("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/phyloseq.rarefied.1k.rds")
seqtab.rare <- data.frame(ps.rare@otu_table)
samdf.rare <- data.frame(ps.rare@sam_data)
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/taxa2.Rdata")

ps.rare <- phyloseq(otu_table(seqtab.rare, taxa_are_rows=FALSE), 
                    sample_data(samdf.rare), 
                    tax_table(taxa2))
ps.rare 
#save(ps.rare,file="/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.rare.Rdata")
#5489 taxa and 167 samples - raw & rarefied

#raw, unrarefied version:
ps.clean = readRDS("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/phyloseq.cleanest.rds")
seqtab.clean <- data.frame(ps.clean@otu_table)
samdf.clean <- data.frame(ps.clean@sam_data)

ps.clean <- phyloseq(otu_table(seqtab.clean, taxa_are_rows=FALSE), 
                    sample_data(samdf.clean), 
                    tax_table(taxa2))
ps.clean
#save(ps.clean,file="/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.clean.Rdata")
#7680 taxa and 172 samples just raw, not rarefied

##TRIMMED VERSIONS:
# trimmed and rarefied
ps.trim.rare = readRDS("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/phyloseq.trim.rare.rds")
seqtab.trim.rare <- data.frame(ps.trim.rare@otu_table)
samdf.trim.rare <- data.frame(ps.trim.rare@sam_data)

ps.trim.rare <- phyloseq(otu_table(seqtab.trim.rare, taxa_are_rows=FALSE), 
                    sample_data(samdf.trim.rare), 
                    tax_table(taxa2))
ps.trim.rare
#save(ps.trim.rare,file="/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.trim.rare.Rdata")
#641 taxa and 165 samples trimmed and rarefied

#checking what happens if we read in the unrarefied version:
ps.trim = readRDS("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/phyloseq.cleanest.trim.rds")
seqtab.trim <- data.frame(ps.trim@otu_table)
samdf.trim <- data.frame(ps.trim@sam_data)

ps.trim <- phyloseq(otu_table(seqtab.trim, taxa_are_rows=FALSE), 
                    sample_data(samdf.trim), 
                    tax_table(taxa2))
ps.trim
#save(ps.trim,file="/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.trim.Rdata")
#641 taxa and 171 samples trimmed but not rarefied
```

```{r read in ps objects}
# can skip to here once you have made the .Rdata files above
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.clean.Rdata")
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.rare.Rdata")
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.trim.rare.Rdata")
load("/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/ps.trim.Rdata")

# remove Control 2, and lineage NA's
ps.clean <- subset_samples(ps.clean,(treat!="Control 2"))
ps.clean <- subset_samples(ps.clean,(!is.na(lineage)))
ps.clean <- subset_samples(ps.clean,(!is.na(treat)))

ps.trim.rare <- subset_samples(ps.trim.rare,(treat!="Control 2"))
ps.trim.rare <- subset_samples(ps.trim.rare,(!is.na(lineage)))
ps.trim.rare <- subset_samples(ps.trim.rare,(!is.na(treat)))

ps.trim <- subset_samples(ps.trim,(treat!="Control 2"))
ps.trim <- subset_samples(ps.trim,(!is.na(lineage)))
ps.trim <- subset_samples(ps.trim,(!is.na(treat)))

ps.rare <- subset_samples(ps.rare,(treat!="Control 2"))
ps.rare <- subset_samples(ps.rare,(!is.na(lineage)))
ps.rare <- subset_samples(ps.rare,(!is.na(treat)))
```

# Diversity

[Notes from phyloseq author](https://rdrr.io/bioc/phyloseq/man/estimate_richness.html)
Visualize alpha-diversity - Should be done on raw, untrimmed dataset

```{r generate div metrics}
df <- data.frame(estimate_richness(ps.clean, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.rare, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.trim, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
#df <- data.frame(estimate_richness(ps.trim.rare, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))

df$frag <- rownames(df)
df.div <- merge(df,samdf,by="frag") #add sample data

#shannon diversity divided by species richness
df.div$even <- df.div$Shannon/(log(df.div$Observed))

str(df.div)
df.div$treat = as.factor(df.div$treat)
df.div$treat = factor(df.div$treat, levels = c("Control", "Low Var","Mod Var","High Var"))

df.div$lineage = as.factor(df.div$lineage)
df.div$sitename = as.factor(df.div$sitename)
```

## Alpha metrics {.tabset}

Looking at everything individually, more of a synthesis below
Alpha diversity is within-sample diversity. Beta diversity, on the other hand, is a measure of similarity or dissimilarity between populations/groups. 

### Shannon
Shannon diversity estimates species diversity, considers number of species (richness) and their relative abundance (evenness)

```{r shannon site}
cols_site_diverging <- c("CI" = "#543005", "PD"= "#bf812d",  "SP"= "#dfc27d",  "BN" = "#003c30", "BS"= "#35978f", "CA"= "#80cdc1")

gg.site.sha <- ggplot(df.div,aes(x=sitename,y=Shannon,color=sitename))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=cols_site_diverging)+
  geom_jitter(alpha=0.5)+
  ylab("Shannon index")+
  xlab("Site")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")
gg.site.sha
```

```{r shannon lineage}
#plotting lineages
cols_lineage <- c("L1" = "#3f007d", "L2" = "#807dba")
df.div.lin = df.div %>%
  drop_na(lineage)

gg.lin.sh <- ggplot(df.div.lin, aes(x=lineage, y=Shannon,color=lineage))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_lineage)+
  geom_jitter(alpha=0.5)+
  xlab("Lineage")+
  ylab("Shannon index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.lin.sh
```

```{r shannon treatment}
#plotting lineages
cols_treat_reds <- c("darkgrey", "#FF9966","#CC3300","#7f0000")

df.div.treat = df.div %>%
  drop_na(treat)

gg.treat.sh <- ggplot(df.div.treat, aes(x=treat, y=Shannon,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Shannon index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")
gg.treat.sh

#facet wrap by lineage
df.div.treat.lin = df.div.treat %>%
  drop_na(lineage)

gg.treat.lin.sh <- ggplot(df.div.treat.lin, aes(x=treat, y=Shannon,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Shannon index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")+
  facet_wrap(~lineage)
gg.treat.lin.sh
```


```{r shannon stats}
#Shannon
shapiro.test(df.div$Shannon) # fine 
leveneTest(df.div$Shannon~treat,data=df.div) #fine

a.div <- aov(Shannon~treat+lineage,data=df.div)
summary(a.div) 
#              Df Sum Sq Mean Sq F value Pr(>F)
# treat         3   2.81  0.9362   1.006  0.392
# lineage       1   0.15  0.1530   0.164  0.686
# Residuals   151 140.51  0.9305               

TukeyHSD(a.div) #ns for raw, rarefied, trimmed, or not
# $treat
#                         diff        lwr       upr     p adj
# Low Var-Control  -0.16026181 -0.7437523 0.4232287 0.8916025
# Mod Var-Control  -0.35255998 -0.9008184 0.1956985 0.3428562
# High Var-Control -0.26280069 -0.8379199 0.3123185 0.6359302
# Mod Var-Low Var  -0.19229818 -0.7570907 0.3724944 0.8129031
# High Var-Low Var -0.10253888 -0.6934411 0.4883633 0.9693824
# High Var-Mod Var  0.08975929 -0.4663806 0.6458992 0.9751104

```

### Simpson

Simpson index considered more of a dominance index, accounts for proportion of species in a sample.
```{r simpson site}
gg.site.sim <- ggplot(df.div,aes(x=sitename,y=InvSimpson,color=sitename))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=cols_site_diverging)+
  geom_jitter(alpha=0.5)+
  ylab("Inv. Simpson index")+
  xlab("Site")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.site.sim
```

```{r simpson lineage}
df.div.lin = df.div %>%
  drop_na(lineage)

gg.lin.sim <- ggplot(df.div.lin, aes(x=lineage, y=InvSimpson,color=lineage))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_lineage)+
  geom_jitter(alpha=0.5)+
  xlab("Lineage")+
  ylab("Inv. Simpson index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")
gg.lin.sim
```

```{r simpson treatment}
#plotting treatment
gg.treat.sim <- ggplot(df.div.treat, aes(x=treat, y=InvSimpson,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Inv. Simpson index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.treat.sim

gg.treat.lin.sim <- ggplot(df.div.treat.lin, aes(x=treat, y=InvSimpson,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Inv. Simpson index")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")+
  facet_wrap(~lineage)
gg.treat.lin.sim
```

```{r simpson stats}
#simpson
shapiro.test(df.div$InvSimpson) #not normal
df.div$si.log <- log(df.div$InvSimpson) 
shapiro.test(df.div$si.log) #fine
leveneTest(df.div$si.log~treat,data=df.div) #fine 

a.div <- aov(si.log~treat+lineage,data=df.div)
summary(a.div) 
#              Df Sum Sq Mean Sq F value Pr(>F)
# treat         3   1.40  0.4651   0.658  0.579
# lineage       1   0.39  0.3872   0.547  0.461
# Residuals   151 106.80  0.7073               

TukeyHSD(a.div) #nothing, rarefied or not
# $treat
#                         diff        lwr       upr     p adj
# Low Var-Control  -0.11727921 -0.6259824 0.3914240 0.9322462
# Mod Var-Control  -0.22011752 -0.6981045 0.2578694 0.6301021
# High Var-Control -0.23424266 -0.7356476 0.2671622 0.6191661
# Mod Var-Low Var  -0.10283832 -0.5952401 0.3895635 0.9484119
# High Var-Low Var -0.11696345 -0.6321284 0.3982015 0.9350293
# High Var-Mod Var -0.01412513 -0.4989833 0.4707331 0.9998439
# 
# $lineage
#            diff        lwr       upr     p adj
# L2-L1 0.1022401 -0.1712182 0.3756985 0.4612308

```

### Richness

```{r richness site}
gg.site.obs <- ggplot(df.div,aes(x=sitename,y=Observed,color=sitename))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=cols_site_diverging)+
  geom_jitter(alpha=0.5)+
  ylab("ASV richness")+
  xlab("Site")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.site.obs
```

```{r richness lineage}
gg.lin.obs <- ggplot(df.div.lin, aes(x=lineage, y=Observed,color=lineage))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_lineage)+
  geom_jitter(alpha=0.5)+
  xlab("Lineage")+
  ylab("ASV richness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.lin.obs
```


```{r richness treatment}
#plotting treatment
gg.treat.obs <- ggplot(df.div.treat, aes(x=treat, y=Observed,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("ASV richness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.treat.obs

gg.treat.lin.obs <- ggplot(df.div.treat.lin, aes(x=treat, y=Observed,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("ASV Richness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")+
  facet_wrap(~lineage)
gg.treat.lin.obs
```

```{r richness stats}
#otu richness
shapiro.test(df.div$Observed) #nope
df.div$obs.log <- log(df.div$Observed)
shapiro.test(df.div$obs.log) #still no...
leveneTest(df.div$obs.log~treat,data=df.div) #fine

a.div <- aov(obs.log~treat+lineage,data=df.div)
summary(a.div) 
#             Df Sum Sq Mean Sq F value Pr(>F)  
# treat         3   6.50  2.1655   2.188 0.0918 .
# lineage       1   0.08  0.0800   0.081 0.7766  
# Residuals   151 149.47  0.9899                 

TukeyHSD(a.div) #no differences
# $treat
#                          diff        lwr        upr     p adj
# Low Var-Control  -0.186595119 -0.7883976 0.41520733 0.8517681
# Mod Var-Control  -0.538856988 -1.1043217 0.02660771 0.0678860
# High Var-Control -0.180503182 -0.7736716 0.41266527 0.8586593
# Mod Var-Low Var  -0.352261868 -0.9347796 0.23025584 0.3982789
# High Var-Low Var  0.006091938 -0.6033548 0.61553871 0.9999937
# High Var-Mod Var  0.358353806 -0.2152397 0.93194731 0.3688171
# 
# $lineage
#             diff        lwr       upr     p adj
# L2-L1 0.04646841 -0.2770364 0.3699732 0.7769487

```

### Evenness

```{r evenness site}
gg.site.eve <- ggplot(df.div,aes(x=sitename,y=even,color=sitename))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=cols_site_diverging)+
  geom_jitter(alpha=0.5)+
  ylab("Evenness")+
  xlab("Site")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.site.eve
```

```{r evenness lineage}
gg.lin.eve <- ggplot(df.div.lin, aes(x=lineage, y=even,color=lineage))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_lineage)+
  geom_jitter(alpha=0.5)+
  xlab("Lineage")+
  ylab("Evenness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")
gg.lin.eve
```

```{r evenness treatment}
#plotting treatment
gg.treat.eve <- ggplot(df.div.treat, aes(x=treat, y=even,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Evenness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none") 
gg.treat.eve

gg.treat.lin.eve <- ggplot(df.div.treat.lin, aes(x=treat, y=even,color=treat))+
  geom_boxplot(outlier.shape=NA)+
  scale_colour_manual(values=cols_treat_reds)+
  geom_jitter(alpha=0.5)+
  xlab("Treatment")+
  ylab("Evenness")+
  theme_bw()+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.title.x = element_blank(), legend.position = "none")+
  facet_wrap(~lineage)
gg.treat.lin.eve
```

```{r evenness stats}
#evenness
shapiro.test(df.div$even) #nope
leveneTest(df.div$even~treat,data=df.div) #fine

a.div <- aov(even~treat+lineage,data=df.div)
summary(a.div) 
#              Df Sum Sq  Mean Sq F value Pr(>F)
# treat         3 0.0496 0.016546   1.134  0.337
# lineage       1 0.0064 0.006379   0.437  0.509
# Residuals   151 2.2033 0.014592               

TukeyHSD(a.div) #no differences, rarefied or not
# $treat
#                          diff         lwr        upr     p adj
# Low Var-Control  -0.019476296 -0.09254259 0.05359000 0.8998382
# Mod Var-Control  -0.005629518 -0.07428396 0.06302492 0.9965708
# High Var-Control -0.046561614 -0.11857963 0.02545641 0.3380400
# Mod Var-Low Var   0.013846778 -0.05687811 0.08457167 0.9569319
# High Var-Low Var -0.027085318 -0.10107973 0.04690909 0.7773362
# High Var-Mod Var -0.040932096 -0.11057347 0.02870928 0.4240158
# 
# $lineage
#             diff         lwr        upr    p adj
# L2-L1 0.01312293 -0.02615457 0.05240043 0.510176

```

## Combine plots so far
```{r evenness stats}
# lineages
plots.lin = ggarrange(gg.lin.sh, gg.lin.sim, gg.lin.obs, gg.lin.eve, ncol=2, nrow=2)
ggsave(plots.lin, filename = "/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/DiversityAnalyses/plots.lineage.pdf", width=8,height=6, units=c("in"), useDingbats=FALSE)

#sites
plots.site = ggarrange(gg.site.sha, gg.site.sim, gg.site.obs, gg.site.eve, ncol=2, nrow=2)
ggsave(plots.site, filename = "/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/DiversityAnalyses/plots.sitename.pdf", width=8,height=6, units=c("in"), useDingbats=FALSE)

#treatment
plots.treat = ggarrange(gg.treat.sh, gg.treat.sim, gg.treat.obs, gg.treat.eve, ncol=2, nrow=2)
ggsave(plots.treat, filename = "/Users/hannahaichelman/Documents/BU/TVE/16S_ITS2/16S_PreStress/DiversityAnalyses/plots.treat.pdf", width=8,height=6, units=c("in"), useDingbats=FALSE)

```

## Hannah stopped here for manuscript. 

## Phylogenetic diversity (Faith's D)

Tutorial from dada2 author [here](https://f1000research.com/articles/5-1492/v2)

```{r packages phylo d}
#install.packages('devtools')
#library(devtools)
#devtools::install_github('twbattaglia/btools')
library(btools)
```

(I'm not running the following chunk every time, because only need to generate the file once)

```{r fasta file generation, eval=FALSE}
# rare.otu <- as.matrix(ps.rare@otu_table)
# rare.taxa <- data.frame(ps.rare@tax_table)
# rownames(rare.taxa)==colnames(rare.otu)
# 
# colnames(rare.otu) <- rare.taxa$V8
# ids <- rownames(rare.taxa)
# 
# path="~/nicfall drive/Moorea_revisions/mr16s_revised/analyze_asv_table/mr16s_rev.cleanest.fasta"
# uniquesToFasta(rare.otu, path, ids = ids, mode = "w", width = 20000)
```

Actual analysis part: 

(I'm not running the following chunk every time, because only need to generate the files once)

```{r phylo d, eval=FALSE}
# seqs <- getSequences("mr16s_rev.cleanest.fasta")
# names(seqs) <- seqs # This propagates to the tip labels of the tree
# saveRDS(seqs,file="phylo.seqs.rev.cleanest.rds")

# #also doing the same thing with a .fasta file post-trimming to see if it makes a difference:
# seqs <- getSequences("mr16s_rev.cleanest.trimmed copy.fasta")
# names(seqs) <- seqs # This propagates to the tip labels of the tree
# saveRDS(seqs,file="phylo.seqs.rev.cleanest.trimmed.rds")
```

Doing this next part in the cluster because it takes forever

```{bash terminal phylo things, eval=FALSE}
##script phylo.R looks like this now:
# library(dada2)
# library(phangorn)
# library(DECIPHER)
#
# seqs <- readRDS("./phylo.seqs.rev.cleanest.rds")
# alignment <- AlignSeqs(DNAStringSet(seqs), anchor=NA)
# phang.align <- phyDat(as(alignment, "matrix"), type="DNA")
# dm <- dist.ml(phang.align)
# treeNJ <- NJ(dm) # Note, tip order != sequence order
# fit = pml(treeNJ, data=phang.align)
#
# ## negative edges length changed to 0!
# 
# fitGTR <- update(fit, k=4, inv=0.2)
# fitGTR <- optim.pml(fitGTR, model="GTR", optInv=TRUE, optGamma=TRUE,
#                     rearrangement = "stochastic", control = pml.control(trace = 0))
# detach("package:phangorn", unload=TRUE)
# saveRDS(fitGTR, file="./phylo.fitgtr.rev.cleanest.rds")

nano phylo.sh

##job file looks like this:
#$ -V # inherit the submission environment
#$ -cwd # start job in submission directory
#$ -N phylo.sh # job name, anything you want
#$ -l h_rt=24:00:00
#$ -M thenicolakriefall@gmail.com
#$ -m be

module load R
module load rstudio
Rscript phylo.R
##exit

##on cluster:
qsub phylo.sh
##saved output as: phylo.fitgtr.rev.cleanest.rds
```

Back in R

```{r back in R}
library(btools)

fitGTR <- readRDS("phylo.fitgtr.rev.cleanest.rds")

#to check for effects of rarefying:
#ps.rare <- ps.clean 

#new phyloseq object:
taxa.rare <- data.frame(ps.rare@tax_table)
seqtab.rare <- data.frame(ps.rare@otu_table)
taxa.rare$sqs <- row.names(taxa.rare) 
taxa.rare$sqs == colnames(seqtab.rare)
row.names(taxa.rare) <- taxa.rare$V8
colnames(seqtab.rare) <- taxa.rare$V8
row.names(taxa.rare) == colnames(seqtab.rare)
taxa.rare <- as.matrix(taxa.rare)

ps.rare.tree <- phyloseq(otu_table(seqtab.rare, taxa_are_rows = FALSE),
                         sample_data(samdf),
                         tax_table(taxa.rare),
                         phy_tree(fitGTR$tree))

pd.div <- estimate_pd(ps.rare.tree)
row.names(df.div) <- df.div$id
df.div.pd <- merge(df.div,pd.div,by=0)

## saving diversity data frame ##
#save & read back in as needed
#write.csv(df.div.pd,file="mr16s_diversity_rev.rare_6.2k.csv") #saving
#df.div <- read.csv("mr16s_diversity_rev.rare_6.2k.csv",row.names=1,header=TRUE) #reading back in

##post trimming:
# fitGTR.trim <- readRDS("phylo.fitgtr.rev.cleanest.trimmed.rds")
# 
# #new phyloseq object:
# taxa.rare <- data.frame(ps.rare.trim@tax_table)
# seqtab.rare <- data.frame(ps.rare.trim@otu_table)
# taxa.rare$sqs <- row.names(taxa.rare) 
# taxa.rare$sqs == colnames(seqtab.rare)
# row.names(taxa.rare) <- taxa.rare$V8
# colnames(seqtab.rare) <- taxa.rare$V8
# row.names(taxa.rare) == colnames(seqtab.rare)
# taxa.rare <- as.matrix(taxa.rare)
# 
# ps.rare.tree <- phyloseq(otu_table(seqtab.rare, taxa_are_rows = FALSE),
#                          sample_data(samdf),
#                          tax_table(taxa.rare),
#                          phy_tree(fitGTR.trim$tree))
# 
# pd.div <- estimate_pd(ps.rare.tree)
# row.names(df.div) <- df.div$id
# df.div.pd <- merge(df.div,pd.div,by=0)
```

### Faith's D Plots {.tabset}

#### Figure - site

```{r faiths d site}
gg.fd.site <- ggplot(df.div.pd,aes(x=site,y=PD,color=site))+
  geom_boxplot(outlier.shape=NA)+
  scale_color_manual(values=c("darkslategray3","darkslategray4","#000004"))+
  geom_jitter(alpha=0.5)+
  ylab("Faith's D")+
  xlab("Site")+
  theme_bw()+
  theme(legend.position="none")
gg.fd.site
```

#### Figure - reef zone

```{r faiths d rz}
gg.fd.rz <- ggplot(df.div.pd, aes(x=zone, y=PD,color=zone,shape=zone))+
  geom_boxplot(outlier.shape=NA)+
  xlab("Reef zone")+
  ylab("Faith's D")+
  theme_bw()+
  scale_shape_manual(values=c(16,15),labels=c("Back reef","Fore reef"))+
  scale_colour_manual(values=c("#ED7953FF","#8405A7FF"),labels=c("Back reef","Fore reef"))+
  #guides(color=guide_legend(title="Reef zone"),shape=guide_legend(title="Reef zone"))+
  theme(legend.position="none")+
  geom_jitter(alpha=0.5)+
  facet_wrap(~site)+
  scale_x_discrete(labels=c("BR","FR"))
gg.fd.rz
```

#### Stats

```{r phylo div stats}
#stats - phylo div
shapiro.test(log(df.div.pd$PD)) #normal
df.div.pd$faith.log <- log(df.div.pd$PD) 
leveneTest(faith.log~site*zone,data=df.div.pd) #fine 

a.div <- aov(faith.log~site/zone,data=df.div.pd)
summary(a.div)
TukeyHSD(a.div) #Tahiti lower than the other two

df.div.pd.mnw <- subset(df.div.pd,site=="MNW")
df.div.pd.mse <- subset(df.div.pd,site=="MSE")
df.div.pd.tnw <- subset(df.div.pd,site=="TNW")

wilcox.test(PD~zone,data=df.div.pd.mnw) #ns
summary(aov(PD~zone,data=df.div.pd.mnw)) #ns
wilcox.test(PD~zone,data=df.div.pd.mse) #ns
summary(aov(PD~zone,data=df.div.pd.mse)) #ns
wilcox.test(PD~zone,data=df.div.pd.tnw) #ns
summary(aov(PD~zone,data=df.div.pd.tnw)) #ns
```

## Synthesizing results {.tabset}

Only significant comparison above is Tahiti being different from Mo'orea NW & SE in ASV richness & phylo diversity (significantly lower)

### Plot by site

```{r plots site}
gg.panels.site <- ggarrange(gg.site.obs,gg.site.eve,gg.site.sha,gg.site.sim,gg.fd.site,nrow=3,ncol=2,labels="AUTO")
gg.panels.site
#ggsave(gg.panels.site,file="bac.div.site.pdf",height=8)
```

### Plot by reef zone

```{r plots rz}
gg.panels.rz <- ggarrange(gg.obs,gg.ev,gg.sh,gg.si,gg.fd.rz,nrow=3,ncol=2,labels="AUTO")
gg.panels.rz
#ggsave(gg.panels.rz,file="bac.div.rz.pdf", height=7)
```







